---
title: "EMT manuscript"
format: html
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE)
library(tidyverse)
library(gt)
library(report)
library(report)
library(patchwork)
here::i_am("web/emt-children.qmd")
source(here::here("scripts/prepare-manuscript-data.R"))

relabel_errors <- function(dat){
  dat <- dat %>% 
    mutate(error_type = case_when(
      error_type == "selection" ~ "Item selection error",
      error_type == "incorrect_placement" ~ "Placement error",
      error_type == "incorrect_object_order" ~ "Object order error", 
      TRUE ~ error_type
    ))
  return(dat)
}
```

Popis vzorku + distribuce veku vs. pohlaví (počty dětí v obtížnostech)

```{r}
df_single <- df_all %>%
  group_by(name, school, age_month, gender) %>%
  slice_head(n = 1) %>%
  ungroup()
```

```{r}
df_vr %>%
  select(difficulty, name, school) %>%
  distinct() %>%
  group_by(difficulty) %>%
  count()
```

```{r}
df_single %>%
  group_by(gender) %>%
  summarise(n = n(), mean_age = mean(age_month), 
            sd_age = sd(age_month), min_age = min(age_month),
            max_age = max(age_month)) %>%
  gt() %>%
  tab_header(title = "Demographics") %>%
  fmt_number(columns = c("mean_age", "sd_age"), decimals = 3) %>%
  tab_spanner(label = "Age", columns = c("mean_age", "sd_age", "min_age", "max_age")) %>%
  # relabel variable names mean and sd and min to Mean, SD, Min, Max
  cols_label(mean_age = "Mean",
              sd_age = "SD",
              min_age = "Min",
              max_age = "Max")
```

A total of `r nrow(df_single)` children participated in the study. Age demographics can be seen in Tab XY. Children were recruited from `r n_distinct(df_single$school)` schools.

All children completed all trials. 

## Average error demographics

```{r}
plt_raw_error <- df_vr %>%
  # relabel fill selection becomes item selection, incorrect_placement becomes placement error 
  relabel_errors() %>%
  select(difficulty, error_type, error_value) %>%
  ggplot(aes(x = as.factor(difficulty), y = error_value)) +
    geom_boxplot() +
    guides(fill = "none") +
    papaja::theme_apa(base_size = 16) +
    geom_jitter(width = 0.2, height = 0.05, alpha = 0.5) +
    labs(x = "", y = "Absolute number of errors") +
    facet_wrap(~error_type)
    

plt_relative_correct <- df_vr %>%
  relabel_errors() %>%
  select(difficulty, relative_correct, error_type) %>%
  ggplot(aes(x = as.factor(difficulty), y = relative_correct)) + 
    geom_jitter(width = 0.2, height = 0.05, alpha = 0.5) +
    labs(x = "Trial difficulty", y = "Relative accuracy (1 is perfect performance)") +
    guides(fill = "none") +
    geom_boxplot() +
    papaja::theme_apa(base_size = 16) +
    facet_wrap(~error_type)

plt_raw_error / plt_relative_correct +
  plot_annotation(
    title = "Error descriptives",
    caption = '',
    theme = theme(plot.title = element_text(size = 26, hjust = 0.5))
  )
```

### Error description

Participant can make one of four distinct errors come in four different types:

- Item selection error: the participant selected the wrong item when given the option at the beginning of the trial
- Incorrect placement: the participant placed the item at a wrong position (not )
- Object order error: the participant placed the item in a wrong order (different from the order in the instruction)
- Location order error: the participant placed the item (correct or incorrect) in a possible location, but different from the order in the instruction


```{r object location order error corr, message = FALSE}
df_location_object_error <- df_vr_unfiltered %>%
  filter(error_type %in% c("incorrect_location_order", "incorrect_object_order")) %>%
  select(-c(relative_error_value, made_error)) %>%
  pivot_wider(names_from = error_type, values_from = error_value)

df_location_object_error %>%
  ggplot(aes(incorrect_location_order, incorrect_object_order)) +
    geom_jitter() +
    geom_smooth(method = "lm") +
    labs(x = "Incorrect location order error", y = "Incorrect object order error", 
        title = "Correlation between incorrect location order error and incorrect object order error")

cor_location_object_order_error <- cor.test(df_location_object_error$incorrect_location_order, 
                        df_location_object_error$incorrect_object_order,
                        method="kendall")
       
```

The object order error and location order errors are in our implementation identical, as it is generally true that when participants places the object in a wrong location, they will also place the object in a wrong order. There are other implementations where the object order error is calculated differently, so the two can be exclusive, but in our case the correlation between object order error and location error is 1 (`r report::report_statistics(cor_location_object_order_error)`), so we left only object order error in the analyses

### Relative accuracy

As there is a difference in the potential number of errors possible at different difficulties (for example, you cannot make four selection mistakes out of three items), we also calculated the **relative accuracy** (1 - errors/maximum posibble errors). So the number 1 in relative accuracy means the best possible performance. This should allow us to consider the performance across different difficulties.

## Increasing number of errors with increasing difficulty
efekt náročnosti je viditelný u všech parametrů (zvyšující se počet chyb) - doplnit statistiku 

```{r}
df_vr %>%
  group_by(error_type) %>%
  group_modify(~report_table(aov(error_value ~ difficulty, data = .x))) %>%
  report_table()
```

::: {.callout-warning}
Just to remind, if one location is incorrect, then more locations will be incorrect after that, as the participant will not be able to place the objects correctly. This will overestimate the number of errors in later trials, as the errors are not independent and it is almost impossible to do "just one error" in the order of objects. Selection and placement are independent from one item to another.
:::

## Effect of age on performance

efekt obtížnosti u jednotlivých podmínek (typ chyby)? - diskutovat, proč nejsou přímo srovnatelné.

efekt věku na výkon
nelze posoudit u dětí (nevyváženo u pohlaví), pravděpodobně nevychází v důsledku nerovnoměrné distribuce věku a pohlaví:
The effect of age months is statistically non-significant and negative (beta = -4.00e-03, 95% CI [-9.72e-03, 1.72e-03], t(416) = -1.37, p = 0.170; Std. beta = -0.07, 95% CI [-0.17, 0.03])
Obrázek korelace s věkem/pohlaví nedávat!
ale lze dokladovat srovnáním děti vs. dospělí u 5 objektů - různé  typy chyb
připravit i graf a skupinovou statistiku srovnávající Relative accuracy
ukázat i všechny náročnosti? děti 3,4,5, dospělí -5,7,9?


korelace se standardními testy:
děti (Sentence repetition a Story Telling):
kumulativní skóry (grafy a statistika) pro:
incorrect_object_order: koreluje se Story r = 0.29	p .002, nesignifikantně s Sentence r=0,16, p .069
incorrect_placement:  koreluje se Story r = 0.18	p .057, nesignifikantně s Sentence r=0,09, p .06321
selection (nemá rozptyl, neuvádět!)

dospělí, zvážit, zda reportovat:
- pokud ano, tak přidat korelaci mezi Office a House
object_order:  OFFICE: RBANS SLOVA: incorrect_order	r = 0.3 (p = < .001)	STORY: r = 0.206 (p = .016)
placement: Office	incorrect_placement	SLOVA r = 0.185 (p = .031)	STORY r = 0.013 (p = .877)
v dospělosti již subtyp nehraje takovou roli, pravděpodobně využívají strategii příběhu i u slov, což u dětí pravděpodobně není



