---
title: "EMT manuscript"
format: html
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE)
library(tidyverse)
library(gt)
library(report)
library(lmerTest)
library(papaja)
library(report)
library(stringr)
library(patchwork)
here::i_am("web/emt-children.qmd")
source(here::here("scripts/prepare-manuscript-data.R"))

relabel_errors <- function(dat){
  dat <- dat %>% 
    mutate(error_type = case_when(
      error_type == "selection" ~ "Item selection",
      error_type == "incorrect_placement" ~ "Placement",
      error_type == "incorrect_object_order" ~ "Object order", 
      TRUE ~ error_type
    ))
  return(dat)
}

# helper to report a single Tukey comparison in APA style inline
report_tukey <- function(tab, comparison) {
  row <- tab[tab[["difficulty comparison"]] == comparison, ]
  str_glue("*M*~diff~ = {round(row$diff, 2)}, 95% CI [{round(row$lwr, 2)}, ",
           "{round(row$upr, 2)}], *p* {papaja::apa_p(row[['p adj']], add_equals = TRUE)}")
}

df_single <- df_all %>%
  group_by(name, school, age_month, gender) %>%
  slice_head(n = 1) %>%
  ungroup()

# adult loading
source(here::here("scripts/loading-adults.R"), encoding = "UTF-8")
df_emt_adults <- df_emt_adults %>%
  filter(Environment == "House")

df_adults_single <- df_emt_adults %>%
  group_by(ID.x) %>%
  slice_head(n=1)

# Combining ---
df_adults_children_combined <- df_emt_adults %>%
  select(error_type, error_value = errors, difficulty,
         relative_correct, ID = bindingid) %>%
  mutate(group = "Adults",
         error_type = case_match(error_type,
          "incorrect_order" ~ "incorrect_object_order", .default = error_type)) %>%
  bind_rows(df_vr %>%
    select(error_type, difficulty, error_value, relative_correct, ID = name) %>%
    mutate(group = "Children"))
```

```{r, eval = FALSE}
df_vr %>%
  select(difficulty, name, school) %>%
  distinct() %>%
  group_by(difficulty) %>%
  count()
```

```{r demographics age}
df_single %>%
  group_by(gender) %>%
  summarise(n = n(), mean_age = mean(age_month), 
            sd_age = sd(age_month), min_age = min(age_month),
            max_age = max(age_month)) %>%
  gt() %>%
  tab_header(title = "Demographics for the children participants") %>%
  fmt_number(columns = c("mean_age", "sd_age"), decimals = 3) %>%
  tab_spanner(label = "Age in months", columns = c("mean_age", "sd_age", "min_age", "max_age")) %>%
  # relabel variable names mean and sd and min to Mean, SD, Min, Max
  cols_label(mean_age = "Mean",
              sd_age = "SD",
              min_age = "Min",
              max_age = "Max")
```

A total of `r nrow(df_single)` children (`r sum(df_single$gender == "female")` girls, `r sum(df_single$gender == "male")` boys) participated in the study. The mean age was `r round(mean(df_single$age_month), 1)` months (*SD* = `r round(sd(df_single$age_month), 1)`, range `r min(df_single$age_month)`–`r max(df_single$age_month)` months). Age demographics by sex are reported in Table XY. Children were recruited from `r n_distinct(df_single$school)` mainstream state kindergartens, and all children completed all trials. As shown in Table XY, age was not balanced across sex (boys were on average older than girls), which constrains the interpretation of age effects within the children sample (see Statistical analysis).

```{r demographics adults}
df_adults_single %>%
  group_by(Gender) %>%
  summarise(n = n(), mean_age = mean(Age), 
            sd_age = sd(Age), min_age = min(Age),
            max_age = max(Age)) %>%
  gt() %>%
  tab_header(title = "Demographics for the adults") %>%
  fmt_number(columns = c("mean_age", "sd_age"), decimals = 3) %>%
  tab_spanner(label = "Age", columns = c("mean_age", "sd_age", "min_age", "max_age")) %>%
  # relabel variable names mean and sd and min to Mean, SD, Min, Max
  cols_label(mean_age = "Mean",
              sd_age = "SD",
              min_age = "Min",
              max_age = "Max")
```

A reference sample of `r nrow(df_adults_single)` healthy adults (range `r min(df_adults_single$Age)`–`r max(df_adults_single$Age)` years) completed the same task in the House environment. Because only the House environment was administered to the HCE group and was always presented as the second session, all adult analyses reported below are restricted to House trials. Demographics are summarised in Table XY; the adult sample was predominantly male, which should be borne in mind when comparing groups.

## Statistical analysis

All analyses were conducted in R (R Core Team, 2023) using the `tidyverse`, `lmerTest`, `report`,  packages. Statistical significance was evaluated at $\alpha$ = .05 (two-tailed), and where multiple tests were performed within a family of comparisons, *p* values were adjusted for the false discovery rate (FDR; Benjamini & Hochberg, 1995). Descriptive statistics are reported as means and standard deviations unless stated otherwise.

```{r citations}
citation("report")
citation("lmerTest")
citation("tidyverse")
```

For each recall trial we derived three error scores — item-selection errors, placement errors, and object-order errors — together with their relative-accuracy transformation (1 − errors / maximum possible errors), which rescales performance onto a common 0–1 metric and allows comparison across trials with different numbers of objects. Because the number of possible errors scales with trial difficulty, all difficulty-spanning and group comparisons were performed on relative accuracy, whereas difficulty-specific analyses are reported on the raw error counts.

Prior to the main analyses, we examined the redundancy of the two order-based error measures. The object-order error and the location-order error were essentially interchangeable in the present implementation: when an item was placed in the wrong location it was, by construction, almost always placed in the wrong order as well. Their association was quantified with Kendall's rank correlation ($\tau$), and because the two indices were perfectly correlated ($\tau$ = 1), only the object-order error was retained in all subsequent analyses to avoid redundancy.

**Effect of task difficulty.** The hypothesis that increasing the number of to-be-remembered objects increases the number of recall errors was tested separately for each error type using one-way analyses of variance (ANOVA) with trial difficulty entered as a factor (`error_value ~ difficulty`). Significant omnibus effects were followed up with Tukey's HSD post-hoc comparisons between difficulty levels. Because errors within a trial are not fully independent (an early order error necessarily propagates to later items), order-error effects across difficulty are interpreted descriptively rather than as evidence of independent error accumulation (see limitations).

**Effect of age and sex in children.** The developmental hypothesis was tested with a linear mixed-effects model predicting relative accuracy from age (in months), sex, and their interaction, with a by-participant random intercept to account for the repeated trials per child (`relative_correct ~ age_months * gender + (1 | name)`). Models were fitted with restricted maximum likelihood, and *p* values for fixed effects were obtained from Satterthwaite-approximated degrees of freedom (`lmerTest`). Because age and sex were unbalanced in the children sample (see Participants), the age × sex interaction is interpreted with caution and age effects should not be read as unconfounded estimates of development.

**Children versus adults.** Differences in episodic-memory performance between preschool children and adults were examined at the highest difficulty level common to interpretation (5 objects) using independent-samples Welch *t*-tests on the raw error counts, conducted separately for placement and object-order errors. The item-selection error was not analysed in this comparison because it showed essentially no variance.

**Convergent validity.** Associations between performance on the virtual task and the standardized neuropsychological measures were tested with Spearman's rank correlations ($\rho$), which are robust to the bounded, non-normal distribution of the accuracy scores. In children, relative accuracy (averaged across trials) for each error type was correlated with the NEPSY Narrative Repetition and Sentence Repetition scores; *p* values were FDR-corrected within each neuropsychological test. In adults, relative accuracy was correlated with the two RBANS list-learning indices (RBANS A.1, word list; RBANS A.2, story); these exploratory correlations are reported uncorrected and interpreted descriptively given the small adult sample.


## Average error demographics

```{r fig.width=10, fig.height=15}
#| out-width: 80%
plt_raw_error <- df_adults_children_combined %>%
  # relabel fill selection becomes item selection, incorrect_placement becomes placement error 
  relabel_errors() %>%
  select(difficulty, error_type, error_value, group) %>%
  ggplot(aes(x = as.factor(difficulty), y = error_value, fill = group)) +
    geom_boxplot() +
    guides(fill = "none") +
    papaja::theme_apa(base_size = 18) +
    geom_jitter(width = 0.2, height = 0.05, alpha = 0.5) +
    labs(x = "", y = "Absolute number of errors") +
    facet_wrap(~error_type)

plt_relative_correct <- df_adults_children_combined %>%
  relabel_errors() %>%
  select(difficulty, relative_correct, error_type, group) %>%
  ggplot(aes(x = as.factor(difficulty), y = relative_correct, fill = group)) + 
    geom_jitter(width = 0.2, height = 0.05, alpha = 0.5) +
    labs(x = "Trial difficulty", y = "Relative accuracy (1 is perfect performance)") +
    geom_boxplot() +
    papaja::theme_apa(base_size = 18) +
    theme(legend.position = "bottom") +
    facet_wrap(~error_type)

plt_raw_error / plt_relative_correct +
  plot_annotation(
    title = "Error descriptives",
    caption = '',
    theme = theme(plot.title = element_text(size = 26, hjust = 0.5))
  )
```

## Increasing number of errors with increasing difficulty

```{r error descriptive table}
df_adults_children_combined %>%
  relabel_errors() %>%
  group_by(error_type, difficulty, group) %>%
  summarise(absolute_errors = str_glue("M = {round(mean(error_value),3)} ({round(sd   (error_value),3)})"),
            relative_accuracy = str_glue("M = {round(mean(relative_correct),3)} ({round(sd(relative_correct),3)})")) %>%
      pivot_longer(cols = c(absolute_errors, relative_accuracy), values_to = "value") %>%
    pivot_wider(names_from = difficulty, values_from = value) %>%
    group_by(name, group) %>%
  gt() %>%
  tab_header(title = "Error descriptives", 
             subtitle = "Mean and standard deviation of absolute errors and relative accuracy across different difficulties")
```

```{r}
df_vr %>%
  group_by(error_type) %>%
  group_modify(~report_table(aov(error_value ~ difficulty, data = .x))) %>%
  gt() %>%
  tab_header(title = "ANOVA for error types and difficulty") %>%
  fmt_number(columns = where(is.numeric), decimals = 3)

aov_order_difficulty <- aov(error_value ~ as.factor(difficulty), data = filter(df_vr, error_type == "incorrect_object_order"))
aov_placement_difficulty <- aov(error_value ~ as.factor(difficulty), data = filter(df_vr, error_type == "incorrect_placement"))
aov_selection_difficulty <- aov(error_value ~ as.factor(difficulty), data = filter(df_vr, error_type == "selection"))
```

We can see that the number of errors increases with increasing difficulty for both object order error (`r report_statistics(aov_order_difficulty)`) and placement error (`r report_statistics(aov_placement_difficulty)`), but not in the selection error (`r report_statistics(aov_selection_difficulty)`).

```{r tukey HSD for difficulty}
tukey_order_difficulty <- TukeyHSD(aov_order_difficulty)[[1]] %>%
  as.data.frame() %>%
  rownames_to_column(var = "difficulty comparison")

tukey_placement_difficulty <- TukeyHSD(aov_placement_difficulty)[[1]] %>%
  as.data.frame() %>%
  rownames_to_column(var = "difficulty comparison")



tukey_order_difficulty %>%
  gt() %>%
    tab_header(title = "Tukey HSD for object order error") %>%
    fmt_number(columns = where(is.numeric), decimals = 3)

tukey_placement_difficulty %>%
  gt() %>%
    tab_header(title = "Tukey HSD for placement errors") %>%
    fmt_number(columns = where(is.numeric), decimals = 3)
```

Tukey HSD post-hoc comparisons indicated that this increase was driven by the most demanding trials. For the object-order error, the 5-object trials produced significantly more errors than both the 3-object trials (`r report_tukey(tukey_order_difficulty, "5-3")`) and the 4-object trials (`r report_tukey(tukey_order_difficulty, "5-4")`), whereas the difference between the 4- and 3-object trials was not significant (`r report_tukey(tukey_order_difficulty, "4-3")`). The same pattern held for placement errors: the 5-object trials yielded significantly more errors than the 3-object (`r report_tukey(tukey_placement_difficulty, "5-3")`) and the 4-object trials (`r report_tukey(tukey_placement_difficulty, "5-4")`), with no significant difference between the 4- and 3-object trials (`r report_tukey(tukey_placement_difficulty, "4-3")`).

::: {.callout-warning}
Just to remind, if one location is incorrect, then more locations will be incorrect after that, as the participant will not be able to place the objects correctly. This will overestimate the number of errors in later trials, as the errors are not independent and it is almost impossible to do "just one error" in the order of objects. Selection and placement are independent from one item to another.
:::

## Effect of age on performance

```{r age and performance, eval = FALSE}
df_vr %>%
  group_by(name, error_type) %>%
  relabel_errors() %>%
  slice_head(n = 1) %>%
  ggplot(aes(age_months, all_trials_avg_relative_correct, color = gender)) +
  geom_point() +
  facet_grid(cols = vars(error_type), scales = "free_y") +
  geom_smooth(method = "lm") +
  labs(x = "Age (months)", y = "Average relative accuracy value (1 - error / difficulty)", 
       title = "Relative accuracy in averaged for all three trials") +
  papaja::theme_apa(base_size = 18) +
  theme(legend.position = "bottom")
```

```{r, results = "asis"}
lmer_rel_age_diff <- lmerTest::lmer(relative_correct ~ age_months*gender + (1|name), data = df_vr)
report(lmer_rel_age_diff)
```

There is no significant effect of age on the performance in the VR task. No effect of gender or the interaction between the two either. 

## Validation with standard tests

Maximum theoretical points is 40 for the story repetition and 34 for the sentence repetition. Higher score is better. 

```{r story and sentence descriptives}
df_all %>%
  group_by(name) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  select(name, NR, SR) %>%
  filter(!is.na(NR)) %>%
  summarise(mean_NR = mean(NR), sd_NR = sd(NR), min_NR = min(NR), max_NR = max(NR),
            mean_SR = mean(SR), sd_SR = sd(SR), min_SR = min(SR), max_SR = max(SR)) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  separate(variable, into = c("var", "test"), sep = "_") %>%
  pivot_wider(names_from = var, values_from = value) %>%
  mutate(test = case_when(
    test == "NR" ~ "Story repetition",
    test == "SR" ~ "Sentence repetition",
    TRUE ~ test
  )) %>%
  gt() %>%
  tab_header(title = "Descriptives of standard tests") %>%
  fmt_number(columns = c(mean, sd), decimals = 3)
```

```{r relationship between sentence, fig.width=10, fig.height=12 }
plt_nr_accuracy <- df_all %>%
  group_by(name, error_type) %>%
  relabel_errors() %>%
  slice_head(n = 1) %>%
  ggplot(aes(all_trials_avg_relative_correct, NR)) +
  geom_point() +
    geom_smooth(method = "lm", color = "brown") +
    facet_wrap(~error_type, scales = "free") +
    labs(x = "Relative accuracy (averaged across all trials)", 
         y = "Story repetition score" , 
         title = "Relationship between relative accuracy and story repetition score for all trials") +
  papaja::theme_apa(base_size = 16)

plt_sr_accuracy <- df_all %>%
  group_by(name, error_type) %>%
  relabel_errors() %>%
  slice_head(n = 1) %>%
  ggplot(aes(all_trials_avg_relative_correct, SR)) +
    geom_point() +
    geom_smooth(method = "lm", color = "violet") +
    facet_wrap(~error_type, scales = "free") +
    labs(x = "Relative accuracy (averaged across all trials)", 
         y = "Sentence repetition score" , 
         title = "Relationship between relative accuracy and sentence repetition score for all trials") +
  papaja::theme_apa(base_size = 16)

plt_nr_accuracy / plt_sr_accuracy 
```

```{r story repetition score}
df_cor_NR <- df_all %>%
  group_by(error_type) %>%
  group_modify(~broom::tidy(cor.test(.$NR, .$relative_correct, method = "spearman"))) %>%
  select(estimate, statistic, p.value) %>%
  mutate(p.value = papaja::apa_p(p.adjust(p.value)),
        value = "Story repetition") %>%
  ungroup() 

df_cor_SR <- df_all %>%
  group_by(error_type) %>%
  group_modify(~broom::tidy(cor.test(.$SR, .$relative_correct, method = "spearman"))) %>%
  select(estimate, statistic, p.value) %>%
  mutate(p.value = papaja::apa_p(p.adjust(p.value)), 
        value = "Sentence repetition") %>%
  ungroup()

df_cor_NR %>%
  bind_rows(df_cor_SR) %>%
  relabel_errors() %>%
  select(value, everything()) %>%
  group_by(value) %>%
  gt() %>%
    tab_header(title = "Correlation between relative accuracy and story/sentence repetition score",
      subtitle = "Spearman correlation") %>%
    fmt_number(columns = where(is.numeric), decimals = 2) %>%
    # add footnote corrected for mutliple comparisons to the header
    tab_footnote(
      footnote = "Corrected for multiple comparisons using FDR method within each category (story or sentence repetitions)",
      locations = cells_column_labels(vars(p.value))
    )
```

## Adults vs. children

```{r comparison adults children descriptives}
df_adults_children_combined %>%
  relabel_errors() %>%
  group_by(group, error_type) %>%
  summarise(report_val = str_glue("{round(mean(error_value), 3)} ({round(sd(error_value), 3)})")) %>%
  pivot_wider(names_from = group, values_from = report_val) %>%
  gt() %>%
    tab_header(title = "Descriptives of error values in adults and children in different error types for 5 item difficulty",
              subtitle = "Mean (SD)")
```

```{r comparison adults children test}
df_adults_children_combined %>%
  # selection error exists only for children (adults have no item-selection
  # measure), so it cannot enter the between-group t-test
  filter(error_type != "selection") %>%
  relabel_errors() %>%
  group_by(error_type) %>%
  group_modify(~report_table(t.test(error_value ~ group, data = .))) %>%
  select(-c(Parameter, Group, starts_with("CI"), Method, Alternative, starts_with("d_CI"))) %>%
  ungroup() %>%
  gt() %>%
    tab_header(title = "Comparison of error values in adults and children in different error types for 5 item difficulty",
              subtitle = "Independent samples t-test") %>%
    fmt_number(columns = where(is.numeric), decimals = 3)
```

### Aduts RBANS

```{r}
df_emt_adults %>%
  # filter(Environment == "House") %>%
  pivot_longer(c(RBANS.A.1, RBANS.A.2), 
              names_to = "RBANS", 
              values_to = "RBANS_value") %>%
  mutate(avg_relative_correct = mean(relative_correct, na.rm = TRUE), .by = ID.x) %>%
  select(avg_relative_correct, Environment, error_type, RBANS, RBANS_value) %>%
  relabel_errors() %>%
  group_by(Environment, error_type, RBANS) %>%
  group_modify(~broom::tidy(cor.test(.$RBANS_value, .$avg_relative_correct, method = "spearman"))) %>%
  summarise(result = str_glue("r = {round(estimate, 3)} (p = {apa_p(p.value)})")) %>%
  pivot_wider(names_from = RBANS, values_from = result) %>%
  ungroup() %>%
  gt() %>%
  tab_header(title = "Correlation between RBANS and EMT for house environment",
             subtitle = "Spearman's rho, not corrected for multiple comparisons. Average house performance across all trials")
```