---
title: "EMT manuscript"
format: html
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE)
library(tidyverse)
library(gt)
library(report)
library(lmerTest)
library(papaja)
library(report)
library(stringr)
library(patchwork)
library(performance)
here::i_am("web/emt-manuscript.qmd")
source(here::here("scripts/prepare-manuscript-data.R"))

relabel_errors <- function(dat){
  dat <- dat %>% 
    mutate(error_type = case_when(
      error_type == "selection" ~ "Item selection",
      error_type == "incorrect_placement" ~ "Placement",
      error_type == "incorrect_object_order" ~ "Object order", 
      TRUE ~ error_type
    ))
  return(dat)
}

# helper to report a single Tukey comparison in APA style inline
report_tukey <- function(tab, comparison) {
  row <- tab[tab[["difficulty comparison"]] == comparison, ]
  str_glue("*M*~diff~ = {round(row$diff, 2)}, 95% CI [{round(row$lwr, 2)}, ",
           "{round(row$upr, 2)}], *p* {papaja::apa_p(row[['p adj']], add_equals = TRUE)}")
}

# helper to report a single fixed-effect term from an lmerTest model in APA style inline
report_lmer_term <- function(model, term) {
  co <- summary(model)$coefficients
  est <- co[term, "Estimate"]; se <- co[term, "Std. Error"]
  df  <- co[term, "df"];       t  <- co[term, "t value"]; p <- co[term, "Pr(>|t|)"]
  ci  <- est + c(-1, 1) * qt(0.975, df) * se
  fmt <- function(x) formatC(x, format = "g", digits = 3)
  str_glue("$\\beta$ = {fmt(est)}, 95% CI [{fmt(ci[1])}, {fmt(ci[2])}], ",
           "*t*({round(df, 0)}) = {round(t, 2)}, *p* {papaja::apa_p(p, add_equals = TRUE)}")
}

# helper to report conditional / marginal R2 of an lmer model inline
report_lmer_r2 <- function(model) {
  r2 <- performance::r2(model)
  str_glue("conditional *R*^2^ = {round(r2$R2_conditional, 3)}, ",
           "marginal *R*^2^ = {round(r2$R2_marginal, 3)}")
}

df_single <- df_all %>%
  group_by(name, school, age_month, gender) %>%
  slice_head(n = 1) %>%
  ungroup()

# adult loading
source(here::here("scripts/loading-adults.R"), encoding = "UTF-8")
df_emt_adults <- df_emt_adults %>%
  filter(Environment == "House")

df_adults_single <- df_emt_adults %>%
  group_by(ID.x) %>%
  slice_head(n=1)

# Combining ---
df_adults_children_combined <- df_emt_adults %>%
  select(error_type, error_value = errors, difficulty,
         relative_correct, ID = bindingid) %>%
  mutate(group = "Adults",
         error_type = case_match(error_type,
          "incorrect_order" ~ "incorrect_object_order", .default = error_type)) %>%
  bind_rows(df_vr %>%
    select(error_type, difficulty, error_value, relative_correct, ID = name) %>%
    mutate(group = "Children"))
```

```{r, eval = FALSE}
df_vr %>%
  select(difficulty, name, school) %>%
  distinct() %>%
  group_by(difficulty) %>%
  count()
```

```{r demographics age}
df_single %>%
  summarise(n = n(), mean_age = mean(age_month), 
            sd_age = sd(age_month), min_age = min(age_month),
            max_age = max(age_month))

df_single %>%
  group_by(gender) %>%
  summarise(n = n(), mean_age = mean(age_month), 
            sd_age = sd(age_month), min_age = min(age_month),
            max_age = max(age_month)) %>%
  gt() %>%
  tab_header(title = "Demographics for the children participants") %>%
  fmt_number(columns = c("mean_age", "sd_age"), decimals = 3) %>%
  tab_spanner(label = "Age in months", columns = c("mean_age", "sd_age", "min_age", "max_age")) %>%
  # relabel variable names mean and sd and min to Mean, SD, Min, Max
  cols_label(mean_age = "Mean",
              sd_age = "SD",
              min_age = "Min",
              max_age = "Max")
```

A total of `r nrow(df_single)` children (`r sum(df_single$gender == "female")` girls, `r sum(df_single$gender == "male")` boys) participated in the study. The mean age was `r round(mean(df_single$age_month), 1)` months (*SD* = `r round(sd(df_single$age_month), 1)`, range `r min(df_single$age_month)`–`r max(df_single$age_month)` months). Age demographics by sex are reported in Table XY. Children were recruited from `r n_distinct(df_single$school)` mainstream state kindergartens, and all children completed all trials. As shown in Table XY, age was not balanced across sex (boys were on average older than girls), which constrains the interpretation of age effects within the children sample (see Statistical analysis).

```{r demographics adults}
df_adults_single %>%
ungroup() %>%
  summarise(n = n(), mean_age = mean(Age), 
            sd_age = sd(Age), min_age = min(Age),
            max_age = max(Age))

df_adults_single %>%
  group_by(Gender) %>%
  summarise(n = n(), mean_age = mean(Age), 
            sd_age = sd(Age), min_age = min(Age),
            max_age = max(Age)) %>%
  gt() %>%
  tab_header(title = "Demographics for the adults") %>%
  fmt_number(columns = c("mean_age", "sd_age"), decimals = 3) %>%
  tab_spanner(label = "Age", columns = c("mean_age", "sd_age", "min_age", "max_age")) %>%
  # relabel variable names mean and sd and min to Mean, SD, Min, Max
  cols_label(mean_age = "Mean",
              sd_age = "SD",
              min_age = "Min",
              max_age = "Max")
```

A reference sample of `r nrow(df_adults_single)` healthy adults (range `r min(df_adults_single$Age)`–`r max(df_adults_single$Age)` years) completed the same task in the House environment. Because only the House environment was administered to the HCE group and was always presented as the second session, all adult analyses reported below are restricted to House trials. Demographics are summarised in Table XY; the adult sample was predominantly male, which should be borne in mind when comparing groups.

## Statistical analysis

All analyses were conducted in R (R Core Team, 2023) using the `tidyverse`, `lmerTest`, `report`,  packages. Statistical significance was evaluated at $\alpha$ = .05 (two-tailed), and where multiple tests were performed within a family of comparisons, *p* values were adjusted for the false discovery rate (FDR; Benjamini & Hochberg, 1995). Descriptive statistics are reported as means and standard deviations unless stated otherwise.

```{r citations}
citation("report")
citation("lmerTest")
citation("tidyverse")
```

For each recall trial we derived three error scores — item-selection errors, placement errors, and object-order errors — together with their relative-accuracy transformation (1 − errors / maximum possible errors), which rescales performance onto a common 0–1 metric and allows comparison across trials with different numbers of objects. Because the number of possible errors scales with trial difficulty, all difficulty-spanning and group comparisons were performed on relative accuracy, whereas difficulty-specific analyses are reported on the raw error counts.

Prior to the main analyses, we examined the redundancy of the two order-based error measures. The object-order error and the location-order error were essentially interchangeable in the present implementation: when an item was placed in the wrong location it was, by construction, almost always placed in the wrong order as well. Their association was quantified with Kendall's rank correlation ($\tau$), and because the two indices were perfectly correlated ($\tau$ = 1), only the object-order error was retained in all subsequent analyses to avoid redundancy.

**Effect of task difficulty.** The hypothesis that increasing the number of to-be-remembered objects increases the number of recall errors was tested separately for each error type using one-way analyses of variance (ANOVA) with trial difficulty entered as a factor (`error_value ~ difficulty`). Significant omnibus effects were followed up with Tukey's HSD post-hoc comparisons between difficulty levels. Because errors within a trial are not fully independent (an early order error necessarily propagates to later items), order-error effects across difficulty are interpreted descriptively rather than as evidence of independent error accumulation (see limitations).

**Effect of age and sex in children.** The developmental hypothesis was tested with a linear mixed-effects model predicting relative accuracy from age (in months), sex, and their interaction, with a by-participant random intercept to account for the repeated trials per child (`relative_correct ~ age_months * gender + (1 | name)`). Models were fitted with restricted maximum likelihood, and *p* values for fixed effects were obtained from Satterthwaite-approximated degrees of freedom (`lmerTest`). Because age and sex were unbalanced in the children sample (see Participants), the age × sex interaction is interpreted with caution and age effects should not be read as unconfounded estimates of development.

**Children versus adults.** Differences in episodic-memory performance between preschool children and adults were examined at the highest difficulty level common to interpretation (5 objects) using independent-samples Welch *t*-tests on the raw error counts, conducted separately for placement and object-order errors. The item-selection error was not analysed in this comparison because it showed essentially no variance.

**Convergent validity.** Associations between performance on the virtual task and the standardized neuropsychological measures were tested with Spearman's rank correlations ($\rho$), which are robust to the bounded, non-normal distribution of the accuracy scores. In children, relative accuracy (averaged across trials) for each error type was correlated with the NEPSY Narrative Repetition and Sentence Repetition scores; *p* values were FDR-corrected within each neuropsychological test. In adults, relative accuracy was correlated with the two RBANS list-learning indices (RBANS A.1, word list; RBANS A.2, story); these exploratory correlations are reported uncorrected and interpreted descriptively given the small adult sample.

## Average error demographics

```{r fig.width=12, fig.height=8}
#| out-width: 80%
plt_raw_error <- df_adults_children_combined %>%
  # relabel fill selection becomes item selection, incorrect_placement becomes placement error 
  relabel_errors() %>%
  select(difficulty, error_type, error_value, group) %>%
  ggplot(aes(x = as.factor(difficulty), y = error_value, fill = group)) +
    geom_boxplot() +
    guides(fill = "none") +
    papaja::theme_apa(base_size = 18) +
    geom_jitter(width = 0.2, height = 0.05, alpha = 0.5) +
    labs(x = "Trial difficulty", title = "Absolute number of errors") +
    facet_wrap(~error_type)

plt_relative_correct <- df_adults_children_combined %>%
  relabel_errors() %>%
  select(difficulty, relative_correct, error_type, group) %>%
  ggplot(aes(x = as.factor(difficulty), y = relative_correct, fill = group)) + 
    geom_jitter(width = 0.2, height = 0.05, alpha = 0.5) +
    labs(x = "Trial difficulty", title = "Relative accuracy (1 is perfect performance)") +
    geom_boxplot() +
    papaja::theme_apa(base_size = 18) +
    theme(legend.position = "bottom") +
    facet_wrap(~error_type)

plt_raw_error / plt_relative_correct +
  plot_annotation(
    caption = '',
  )
```

## Increasing number of errors with increasing difficulty

```{r error descriptive table}
df_adults_children_combined %>%
  relabel_errors() %>%
  group_by(error_type, difficulty, group) %>%
  summarise(absolute_errors = str_glue("M = {round(mean(error_value),3)} ({round(sd   (error_value),3)})"),
            relative_accuracy = str_glue("M = {round(mean(relative_correct),3)} ({round(sd(relative_correct),3)})")) %>%
      pivot_longer(cols = c(absolute_errors, relative_accuracy), values_to = "value") %>%
    pivot_wider(names_from = difficulty, values_from = value) %>%
    group_by(name, group) %>%
  gt() %>%
  tab_header(title = "Error descriptives", 
             subtitle = "Mean and standard deviation of absolute errors and relative accuracy across different difficulties")
```

```{r}
df_vr %>%
  group_by(error_type) %>%
  group_modify(~report_table(aov(error_value ~ difficulty, data = .x))) %>%
  gt() %>%
  tab_header(title = "ANOVA for error types and difficulty") %>%
  fmt_number(columns = where(is.numeric), decimals = 3)

aov_order_difficulty <- aov(error_value ~ as.factor(difficulty), data = filter(df_vr, error_type == "incorrect_object_order"))
aov_placement_difficulty <- aov(error_value ~ as.factor(difficulty), data = filter(df_vr, error_type == "incorrect_placement"))
aov_selection_difficulty <- aov(error_value ~ as.factor(difficulty), data = filter(df_vr, error_type == "selection"))
```

### Task difficulty and errors

We used analysis of variance to test whether the number of errors (raw values) differed between conditions. As expected, the number of errors increased with increasing difficulty for both the object-order error (`r report_statistics(aov_order_difficulty)`) and the placement error (`r report_statistics(aov_placement_difficulty)`), but, interestingly, not for the selection error (`r report_statistics(aov_selection_difficulty)`). This last result is most likely explained by the fact that the children made almost no selection mistakes regardless of difficulty (see Table XY).

```{r tukey HSD for difficulty}
tukey_order_difficulty <- TukeyHSD(aov_order_difficulty)[[1]] %>%
  as.data.frame() %>%
  rownames_to_column(var = "difficulty comparison")

tukey_placement_difficulty <- TukeyHSD(aov_placement_difficulty)[[1]] %>%
  as.data.frame() %>%
  rownames_to_column(var = "difficulty comparison")

tukey_order_difficulty %>%
  gt() %>%
    tab_header(title = "Tukey HSD for object order error") %>%
    fmt_number(columns = where(is.numeric), decimals = 3)

tukey_placement_difficulty %>%
  gt() %>%
    tab_header(title = "Tukey HSD for placement errors") %>%
    fmt_number(columns = where(is.numeric), decimals = 3)
```

Tukey HSD post-hoc comparisons indicated that this increase was driven by the most demanding trials. For the object-order error, the 5-object trials produced significantly more errors than both the 3-object trials (`r report_tukey(tukey_order_difficulty, "5-3")`) and the 4-object trials (`r report_tukey(tukey_order_difficulty, "5-4")`), whereas the difference between the 4- and 3-object trials was not significant (`r report_tukey(tukey_order_difficulty, "4-3")`). The same pattern held for placement errors: the 5-object trials yielded significantly more errors than the 3-object (`r report_tukey(tukey_placement_difficulty, "5-3")`) and the 4-object trials (`r report_tukey(tukey_placement_difficulty, "5-4")`), with no significant difference between the 4- and 3-object trials (`r report_tukey(tukey_placement_difficulty, "4-3")`).

::: {.callout-warning}
Just to remind, if one location is incorrect, then more locations will be incorrect after that, as the participant will not be able to place the objects correctly. This will overestimate the number of errors in later trials, as the errors are not independent and it is almost impossible to do "just one error" in the order of objects. Selection and placement are independent from one item to another.
:::

## Effect of age on performance

Our general developmental hypothesis stated that episodic memory performance would improve with increasing age among preschool children. To examine the effect of the child's age on overall performance, we fitted, separately for each error type, a linear mixed model predicting relative accuracy from the child's age and task difficulty (formula: `relative accuracy ~ age_months + difficulty`), with the participant entered as a random effect (`~ 1 | name`). The relationship between age, difficulty, and relative accuracy is illustrated in Figure XY.

```{r age difficulty relative accuracy plot, fig.width=10, fig.height=8}
#| fig-cap: "Relative accuracy as a function of age (in months) across error types and trial difficulties. Each point is a single trial; lines show ordinary least-squares fits with 95% confidence bands, separately for girls and boys. A relative accuracy of 1 (dashed line) denotes errorless performance."
df_vr %>%
  filter(error_type != "selection") %>%
  relabel_errors() %>%
  mutate(difficulty = factor(difficulty, labels = paste(sort(unique(difficulty)), "objects"))) %>%
  ggplot(aes(age_months, relative_correct, colour = gender, fill = gender)) +
    geom_hline(yintercept = 1, linetype = "dashed", colour = "grey50") +
    geom_point(alpha = 0.4, size = 1.6) +
    geom_smooth(method = "lm", alpha = 0.15) +
    facet_grid(rows = vars(difficulty), cols = vars(error_type)) +
    scale_colour_manual(values = c(female = "#1B9E77", male = "#7570B3"),
                        aesthetics = c("colour", "fill"),
                        labels = c(female = "Girls", male = "Boys"),
                        name = NULL) +
    labs(x = "Age (months)",
         y = "Relative accuracy (1 = errorless performance)") +
    papaja::theme_apa(base_size = 14) +
    theme(legend.position = "bottom",
          panel.spacing = unit(1, "lines"))
```

```{r age and performance, eval = FALSE}
df_vr %>%
  group_by(name, error_type) %>%
  relabel_errors() %>%
  slice_head(n = 1) %>%
  ggplot(aes(age_months, all_trials_avg_relative_correct, color = gender)) +
  geom_point() +
  facet_grid(cols = vars(error_type), scales = "free_y") +
  geom_smooth(method = "lm") +
  labs(x = "Age (months)", y = "Average relative accuracy value (1 - error / difficulty)", 
       title = "Relative accuracy in averaged for all three trials") +
  papaja::theme_apa(base_size = 18) +
  theme(legend.position = "bottom")
```

### Selection accuracy and age
```{r age selection error lmer}
lmer_rele_age_diff_selection <- lmerTest::lmer(relative_correct ~ age_months + difficulty + (1|name), data = filter(df_vr, error_type == "selection"))
report(lmer_rele_age_diff_selection)
report_table(lmer_rele_age_diff_selection) %>%
  gt() %>%
  tab_header(title = "Linear mixed-effects model for selection error with age and difficulty as fixed effects and participant as random effect") %>%
  fmt_number(columns = where(is.numeric), decimals = 3)
```

For item-selection accuracy, the model's total explanatory power was weak (`r report_lmer_r2(lmer_rele_age_diff_selection)`). The effect of age was statistically non-significant and negative (`r report_lmer_term(lmer_rele_age_diff_selection, "age_months")`), and the effect of difficulty was likewise non-significant and negative (`r report_lmer_term(lmer_rele_age_diff_selection, "difficulty")`). This is consistent with the near-absence of selection errors reported above.

### Object order accuracy and age
```{r age order error lmer}
lmer_rele_age_diff_order <- lmerTest::lmer(relative_correct ~ age_months + difficulty + (1|name), data = filter(df_vr, error_type == "incorrect_object_order"))
report(lmer_rele_age_diff_order)
report_table(lmer_rele_age_diff_order) %>%
  gt() %>%
  tab_header(title = "Linear mixed-effects model for object order error with age and difficulty as fixed effects and participant as random effect") %>%
  fmt_number(columns = where(is.numeric), decimals = 3)
```

The model for object-order accuracy showed moderate explanatory power (`r report_lmer_r2(lmer_rele_age_diff_order)`). The effect of difficulty was statistically significant and negative (`r report_lmer_term(lmer_rele_age_diff_order, "difficulty")`), mirroring the ANOVA results reported above, whereas the effect of age was non-significant, although also negative (`r report_lmer_term(lmer_rele_age_diff_order, "age_months")`).

### Placement accuracy and age
```{r age placement accuracy lmer results}
lmer_rele_age_diff_placement <- lmerTest::lmer(relative_correct ~ age_months + difficulty + (1|name), data = filter(df_vr, error_type == "incorrect_placement"))
report(lmer_rele_age_diff_placement)
report_table(lmer_rele_age_diff_placement) %>%
  gt() %>%
  tab_header(title = "Linear mixed-effects model for placement error with age and difficulty as fixed effects and participant as random effect") %>%
  fmt_number(columns = where(is.numeric), decimals = 3)
```

Finally, the model for placement accuracy had substantial explanatory power (`r report_lmer_r2(lmer_rele_age_diff_placement)`). As with object-order accuracy, the effect of difficulty was statistically significant and negative (`r report_lmer_term(lmer_rele_age_diff_placement, "difficulty")`), while the effect of age was statistically non-significant and negative (`r report_lmer_term(lmer_rele_age_diff_placement, "age_months")`).

Taken together, these models indicate that task difficulty was a robust predictor of object-order and placement accuracy, whereas the child's age within this narrow preschool range was not significantly associated with performance in any of the three error types. We return to possible explanations for this null age effect in the Discussion.

## Validation with standard tests

To evaluate the convergent validity of the virtual task, we examined whether performance in the virtual house related to the children's performance on two standardized verbal-memory measures from the NEPSY battery: the Narrative (story) Repetition and the Sentence Repetition subtests. We hypothesized that children who performed better in the virtual episodic memory task would also score higher on these established memory measures. The maximum theoretical score is 40 points for the story repetition and 34 points for the sentence repetition, with higher scores indicating better performance; descriptive statistics for both subtests are reported in Table XY.

Because the relative-accuracy scores and the test scores were bounded and not normally distributed, the associations were quantified with Spearman's rank correlations ($\rho$), computed separately for each error type and corrected for multiple comparisons within each standardized test using the false discovery rate. The relationships are displayed in Figure XY and summarized in Table XY.

```{r story and sentence descriptives}
df_all %>%
  group_by(name) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  select(name, NR, SR) %>%
  filter(!is.na(NR)) %>%
  summarise(mean_NR = mean(NR), sd_NR = sd(NR), min_NR = min(NR), max_NR = max(NR),
            mean_SR = mean(SR), sd_SR = sd(SR), min_SR = min(SR), max_SR = max(SR)) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  separate(variable, into = c("var", "test"), sep = "_") %>%
  pivot_wider(names_from = var, values_from = value) %>%
  mutate(test = case_when(
    test == "NR" ~ "Story repetition",
    test == "SR" ~ "Sentence repetition",
    TRUE ~ test
  )) %>%
  gt() %>%
  tab_header(title = "Descriptives of standard tests") %>%
  fmt_number(columns = c(mean, sd), decimals = 3)
```

```{r relationship between sentence, fig.width=12, fig.height=8}
plt_nr_accuracy <- df_all %>%
  group_by(name, error_type) %>%
  relabel_errors() %>%
  slice_head(n = 1) %>%
  ggplot(aes(all_trials_avg_relative_correct, NR)) +
  geom_point() +
    geom_smooth(method = "lm", color = "brown") +
    facet_wrap(~error_type, scales = "free") +
    labs(x = "Relative accuracy (averaged across all trials)", 
         y = "Story repetition score" , 
         title = "Relationship between relative accuracy and story repetition score for all trials") +
  papaja::theme_apa(base_size = 16)

plt_sr_accuracy <- df_all %>%
  group_by(name, error_type) %>%
  relabel_errors() %>%
  slice_head(n = 1) %>%
  ggplot(aes(all_trials_avg_relative_correct, SR)) +
    geom_point() +
    geom_smooth(method = "lm", color = "violet") +
    facet_wrap(~error_type, scales = "free") +
    labs(x = "Relative accuracy (averaged across all trials)", 
         y = "Sentence repetition score" , 
         title = "Relationship between relative accuracy and sentence repetition score for all trials") +
  papaja::theme_apa(base_size = 16)

plt_nr_accuracy / plt_sr_accuracy 
```

```{r story repetition score}
df_cor_NR <- df_all %>%
  group_by(error_type) %>%
  group_modify(~broom::tidy(cor.test(.$NR, .$relative_correct, method = "spearman"))) %>%
  select(estimate, statistic, p.value) %>%
  ungroup() %>%
  mutate(p.value = papaja::apa_p(p.adjust(p.value, method = "fdr")),
        value = "Story repetition") 

df_cor_SR <- df_all %>%
  group_by(error_type) %>%
  group_modify(~broom::tidy(cor.test(.$SR, .$relative_correct, method = "spearman"))) %>%
  select(estimate, statistic, p.value) %>%
  ungroup() %>%
  mutate(p.value = papaja::apa_p(p.adjust(p.value, method = "fdr")), 
        value = "Sentence repetition") 

df_cor_NR %>%
  bind_rows(df_cor_SR) %>%
  relabel_errors() %>%
  select(value, everything()) %>%
  group_by(value) %>%
  gt() %>%
    tab_header(title = "Correlation between relative accuracy and story/sentence repetition score",
      subtitle = "Spearman correlation") %>%
    fmt_number(columns = where(is.numeric), decimals = 2) %>%
    # add footnote corrected for mutliple comparisons to the header
    tab_footnote(
      footnote = "Corrected for multiple comparisons using FDR method within each category (story or sentence repetitions)",
      locations = cells_column_labels(vars(p.value))
    )

# helper to report a single Spearman correlation from the df_cor tables inline.
# p.value is already an apa_p() string (e.g. ".002" or "< .001"), so prepend
# "= " unless it is an inequality.
report_validation_cor <- function(tab, et) {
  row <- tab[tab$error_type == et, ]
  p_str <- if (startsWith(trimws(row$p.value), "<")) row$p.value else paste("=", row$p.value)
  str_glue("$\\rho$ = {round(row$estimate, 2)}, *p* {p_str}")
}
```

The pattern of associations differed across the two standardized tests. Performance on the **story repetition** subtest was positively and significantly associated with object-order accuracy in the virtual task (`r report_validation_cor(df_cor_NR, "incorrect_object_order")`), with the relationship to placement accuracy being positive but not reaching significance after correction (`r report_validation_cor(df_cor_NR, "incorrect_placement")`). In contrast, correlations with the **sentence repetition** subtest were generally weaker and did not reach significance for either object-order accuracy (`r report_validation_cor(df_cor_SR, "incorrect_object_order")`) or placement accuracy (`r report_validation_cor(df_cor_SR, "incorrect_placement")`). The item-selection error showed almost no variance in the children sample and is therefore not interpreted here. Overall, these results provide preliminary support for the convergent validity of the virtual task — in particular, the object-order component, which captures the temporal sequencing aspect of episodic memory, tracked the narrative-memory measure most closely.

## Adults vs. children

We hypothesized that healthy adults would outperform preschool children, reflecting the still-immature state of episodic memory in early childhood. Because the two groups completed different difficulty levels (children were tested with 2–5 objects and adults with 3–11 objects), a direct comparison of the raw number of errors across all trials would confound group with task difficulty. We therefore restricted the between-group comparison to the single difficulty level common to both samples — the 5-object trials — and compared the raw number of errors with independent-samples Welch *t*-tests, conducted separately for the placement and object-order errors. The item-selection error was excluded because it is recorded only in the children's version of the task and showed virtually no variance.

```{r adults children difficulty 5 data}
df_adults_children_5 <- df_adults_children_combined %>%
  filter(difficulty == 5, error_type != "selection")
```

```{r comparison adults children descriptives}
df_adults_children_5 %>%
  relabel_errors() %>%
  group_by(group, error_type) %>%
  summarise(report_val = str_glue("{round(mean(error_value), 3)} ({round(sd(error_value), 3)})")) %>%
  pivot_wider(names_from = group, values_from = report_val) %>%
  gt() %>%
    tab_header(title = "Descriptives of error values in adults and children for the 5-object trials",
              subtitle = "Mean (SD)")
```

```{r comparison adults children test}
tbl_adults_children_test <- df_adults_children_5 %>%
  relabel_errors() %>%
  group_by(error_type) %>%
  group_modify(~report_table(t.test(error_value ~ group, data = .))) %>%
  ungroup()

tbl_adults_children_test %>%
  select(-c(Parameter, Group, starts_with("CI"), Method, Alternative, starts_with("d_CI"))) %>%
  gt() %>%
    tab_header(title = "Comparison of error values in adults and children for the 5-object trials",
              subtitle = "Independent samples t-test") %>%
    fmt_number(columns = where(is.numeric), decimals = 3)
```

As predicted, adults made significantly fewer errors than children on the 5-object trials for both error types. For the object-order error, children (`r df_adults_children_5 %>% filter(group == "Children", error_type == "incorrect_object_order") %>% summarise(s = str_glue("*M* = {round(mean(error_value),2)}, *SD* = {round(sd(error_value),2)}")) %>% pull(s)`) made substantially more errors than adults (`r df_adults_children_5 %>% filter(group == "Adults", error_type == "incorrect_object_order") %>% summarise(s = str_glue("*M* = {round(mean(error_value),2)}, *SD* = {round(sd(error_value),2)}")) %>% pull(s)`), `r tbl_adults_children_test %>% filter(error_type == "Object order") %>% summarise(s = str_glue("*t*({round(df_error,1)}) = {round(t,2)}, *p* {papaja::apa_p(p, add_equals = TRUE)}")) %>% pull(s)`. The same was true for the placement error, with children (`r df_adults_children_5 %>% filter(group == "Children", error_type == "incorrect_placement") %>% summarise(s = str_glue("*M* = {round(mean(error_value),2)}, *SD* = {round(sd(error_value),2)}")) %>% pull(s)`) making more errors than adults (`r df_adults_children_5 %>% filter(group == "Adults", error_type == "incorrect_placement") %>% summarise(s = str_glue("*M* = {round(mean(error_value),2)}, *SD* = {round(sd(error_value),2)}")) %>% pull(s)`), `r tbl_adults_children_test %>% filter(error_type == "Placement") %>% summarise(s = str_glue("*t*({round(df_error,1)}) = {round(t,2)}, *p* {papaja::apa_p(p, add_equals = TRUE)}")) %>% pull(s)`. These results confirm that, even at the difficulty level the children found most demanding, episodic memory performance in preschoolers remained well below adult levels, consistent with the protracted development of episodic memory through childhood.

### Adult convergent validity (RBANS)

As a parallel check on convergent validity in the adult sample, we correlated each participant's average relative accuracy in the house task with their scores on the two list-learning indices of the RBANS (A.1, immediate word-list recall; A.2, story recall), using Spearman's rank correlations. As shown in Table XY, average task accuracy was positively and significantly associated with both RBANS indices (word list: `r df_emt_adults %>% pivot_longer(c(RBANS.A.1, RBANS.A.2), names_to = "RBANS", values_to = "v") %>% mutate(acc = mean(relative_correct, na.rm = TRUE), .by = ID.x) %>% filter(RBANS == "RBANS.A.1") %>% summarise(s = {ct <- suppressWarnings(cor.test(v, acc, method = "spearman")); str_glue("$\\rho$ = {round(ct$estimate,2)}, *p* {papaja::apa_p(ct$p.value, add_equals = TRUE)}")}) %>% pull(s)`; story: `r df_emt_adults %>% pivot_longer(c(RBANS.A.1, RBANS.A.2), names_to = "RBANS", values_to = "v") %>% mutate(acc = mean(relative_correct, na.rm = TRUE), .by = ID.x) %>% filter(RBANS == "RBANS.A.2") %>% summarise(s = {ct <- suppressWarnings(cor.test(v, acc, method = "spearman")); str_glue("$\\rho$ = {round(ct$estimate,2)}, *p* {papaja::apa_p(ct$p.value, add_equals = TRUE)}")}) %>% pull(s)`), mirroring the convergent-validity pattern observed in children and supporting the interpretation of the virtual task as a measure of episodic memory.

```{r}
df_emt_adults %>%
  # filter(Environment == "House") %>%
  pivot_longer(c(RBANS.A.1, RBANS.A.2), 
              names_to = "RBANS", 
              values_to = "RBANS_value") %>%
  mutate(avg_relative_correct = mean(relative_correct, na.rm = TRUE), .by = ID.x) %>%
  select(avg_relative_correct, Environment, error_type, RBANS, RBANS_value) %>%
  relabel_errors() %>%
  group_by(Environment, error_type, RBANS) %>%
  group_modify(~broom::tidy(cor.test(.$RBANS_value, .$avg_relative_correct, method = "spearman"))) %>%
  summarise(result = str_glue("r = {round(estimate, 3)} (p = {apa_p(p.value)})")) %>%
  pivot_wider(names_from = RBANS, values_from = result) %>%
  ungroup() %>%
  gt() %>%
  tab_header(title = "Correlation between RBANS and EMT for house environment",
             subtitle = "Spearman's rho, not corrected for multiple comparisons. Average house performance across all trials")
```